---
title: "GraphLD-style precision operators"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GraphLD-style precision operators}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(RcppLDGM)
library(Matrix)
```

GraphLD represents LDGM precision blocks as sparse linear operators. RcppLDGM
keeps the same zero-based indexing convention at the file/API boundary while
returning ordinary R vectors, matrices, sparse matrices, and data frames.

```{r precision}
edges <- ldgm_edge_list(
  from = c(0L, 1L, 2L, 0L, 1L),
  to = c(0L, 1L, 2L, 1L, 2L),
  weight = c(2, 3, 4, 0.1, 0.2)
)
P <- ldgm_sparse_precision(edges)
P

x <- c(1, 2, 3)
ldgm_precision_multiply(P, x)
ldgm_precision_solve(P, x)
ldgm_inverse_diagonal(P, method = "exact")
```

An `ldgm_precision` object stores sparse precision data plus variant metadata.
Selected views use GraphLD-style Schur-complement semantics.

```{r selected-view}
variant_info <- data.frame(
  index = c(0L, 0L, 1L, 2L),
  SNP = c("rs1", "rs1_proxy", "rs2", "rs3")
)
block <- ldgm_precision(P, variant_info)
selected <- ldgm_precision_select(block, c(0L, 2L))
ldgm_precision_matrix(selected)
ldgm_precision_multiply(selected, c(1, 1))
```

The R API is functional: update helpers return a modified copy instead of
mutating in place.

```{r updates}
scaled <- ldgm_precision_scale(selected, 2)
updated <- ldgm_precision_update_element(selected, index = 1L, value = 0.5)
ldgm_precision_multiply(scaled, c(1, 1))
ldgm_precision_multiply(updated, c(1, 1))
```

When multiple variants share the same precision index, `ldgm_variant_solve()`
sums right-hand-side values by index, solves once, and broadcasts the solution
back to variant rows. This mirrors GraphLD's `variant_solve()` behavior.

```{r variant-solve}
ldgm_variant_solve(block, c(1, 2, 3, 4))
```

A small deterministic graphREML block can be evaluated with exact inverse
diagonals for testing and examples.

```{r reml-block}
z <- c(1.2, -0.4, 0.7)
annotations <- cbind(baseline = 1, coding = c(0, 1, 0))
fit <- ldgm_reml_block(
  P,
  z,
  annotations,
  params = c(0, 0),
  sample_size = 1000,
  diagonal_method = "exact"
)
fit$likelihood
fit$gradient
fit$hessian
```
