---
title: "Table-backed LDGM construction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Table-backed LDGM construction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(RcppLDGM)
```

RcppLDGM ports the upstream LDGM construction pipeline around a narrow,
table-first boundary. The native kernels consume canonical tree-diff tables. In
production these tables can be extracted from a `.trees` file with
`ldgm_tree_tables_from_tskit()`, which uses the vendored tskit C API.

This vignette uses a tiny in-memory table bundle so it can run during package
checks without Python or external data.

```{r toy-tables}
initial_edges <- data.frame(
  left = c(0, 0),
  right = c(1, 1),
  parent = c(2L, 2L),
  child = c(0L, 1L)
)
empty_transitions <- data.frame(transition = integer(), left = numeric())
empty_edges_out <- data.frame(transition = integer(), child = integer())
empty_edges_in <- data.frame(
  transition = integer(), left = numeric(), right = numeric(),
  parent = integer(), child = integer()
)
empty_node_state <- data.frame(
  transition = integer(), node = integer(), prev_parent = integer(),
  curr_parent = integer(), time = numeric(), curr_num_samples = integer()
)
mutations <- data.frame(
  mutation = c(0L, 1L),
  site = c(0L, 1L),
  position = c(0.1, 0.2),
  node = c(0L, 1L),
  ancestral_state = c("A", "C"),
  derived_state = c("G", "T")
)

tables <- ldgm_tree_tables(
  initial_edges = initial_edges,
  transitions = empty_transitions,
  edges_out = empty_edges_out,
  edges_in = empty_edges_in,
  node_state = empty_node_state,
  sample_nodes = 0:1,
  mutations = mutations
)
tables
```

`ldgm_brick_ts()` exposes the upstream-style bricking entry point.

```{r brick-ts}
ldgm_brick_ts(tables)
```

`ldgm_make_ldgm()` runs bricking, mutation mapping, brick-haplotype graph
construction, reduction, and SNP-list generation.

```{r make-ldgm}
result <- ldgm_make_ldgm(tables, path_threshold = 4, return_intermediates = TRUE)
result$bricked_edges
result$bricks_to_muts
result$snplist
ldgm_return_edgelist(result$graph)
```

For real conformance, use goldens generated from pinned upstream Python code:

```{sh, eval = FALSE}
make upstream-python
make upstream-ldgm-conformance
```
